Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-05-10 papers/paper_alt_theor_bounds/arxiv_v1/paper_alt_theor_bounds.tex
hs: v1 of arXiv

\documentclass[twocolumn,
               prd,
               aps,
               superscriptaddress,
               tightenlines,
               nofootinbib,
               eqsecnum,
               amsfonts,
               amsmath,
               longbibliography]{revtex4-1}
\usepackage[flushleft]{threeparttable}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
% \usepackage{newtxtext}
% \usepackage{newtxmath}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}

@@

\begin{figure}[htb]
\includegraphics[width=\columnwidth]{figs/example_waveform_cubicEFT.pdf}
\caption{(Color online). Gravitational-wave signal with GW150914-like parameters
for both GR (solid line) and cubic EFT of GR (dashed line) with the leading-order
$n=0$ modifications to the fundamental QNM, for $\ell_{\rm cEFT} = 65$~km.
%
The former is computed with the \SEOB{} model, while the latter with the
\pSEOB{} model, with ringdown modifications according to the results in
Table~\ref{tab:ref_theories_qnms}.
%
Top panel: the $+$ polarization $h_{+}(t)$. Bottom panel: the GW amplitude
$|\tilde{h}(t)|$ (left axes) and the instantaneous frequency $f(t)$ (right axes).
}
\label{fig:example_waveform}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function on coupling constant $\ell_{\rm EdGB}$ in EdGB gravity
for GW150914 (top panel), GW200129 (middle panel)
and from combining results (bottom panel).
%
In all panels, different line colors correspond to the inclusion ($N_{\rm max} = 1$) or not ($N_{\rm max} = 0$)
of the linear-in-spin QNM correction.
%
The joint posteriors are shown for illustrative purposes only. As we explain in Fig.~\ref{fig:edgb_cdf} and
in the main text, our analysis of these events fails to satisfy the EFT bound~\eqref{eq:eft_bound}. We mark
with vertical lines the $90\%$ upper credible intervals.}
\label{fig:edgb_exec_sum}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\columnwidth]{figs/edgb_gamma.pdf}
\caption{(Color online).~Posterior distribution function on the dimensionless parameter $\gamma_{\rm EdGB}$,
defined in Eq.~\eqref{eq:def_gamma_gb}. Different line colors distinguish between events, while
different line styles distinguish between different $N_{\rm max}$.
%
We see that this parameter which controls the ParSpec expansion in EdGB gravity
does have maximum support at $\gamma_{\rm EdGB} = 0$. This shows that our model is consistent with GR.
The right panel shows the PDFs in the range $0 \leqslant \gamma_{\rm EdGB} \leqslant 1$. Note how
the curves are flat in this range for both events and when $N_{\rm max} = 1$.}
\label{fig:gamma_gb}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ for EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{\rm f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines mark the $90\%$ credible levels. Having set the maximum value
of $\varepsilon$ to be 1/2, we see that in no situation the curves pass
through the 0.90 lines. This means that no bound on $\ell_{\rm EdGB}$ can be placed
with the events we analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_posteriors_combined.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for dCS gravity.
%
We stress that posteriors obtained including $n=1$ corrections, violate conditions~\eqref{eq:parspec_bound}
and therefore should not be used to draw meaningful conclusions. We show them for illustrative purposes
and also to emphasize the importance of taking conditions~\eqref{eq:eft_bound} and~\eqref{eq:parspec_bound}
simultaneously into consideration when analyzing the results of the parameter estimation.
}
\label{fig:dCS_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_cdf_varying_threshold.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_cdf}, but for dCS gravity.
%
We see that the CDF curves for $\gm = M_{\rm f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $n = 1$ dCS corrections to
the dominant QNM.
}
\label{fig:dcs_cdf}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=0.9\columnwidth]{figs/tmp_GW150914_intrinsic_params_remnant.pdf}
\caption{(Color online).~Corner plot showing that the inferred final spin $\chi_{\rm f}$,
remnant mass $M_{\rm f}$ for GW150914, using the same waveform model,
but without (purple contours) and with the non-GR parameters
different from zero (blue contours), for
dCS gravity and $N_{\rm max}=0$.
%
The contours represent 90\% credible levels.
%
We see that the introduction of the non-GR parameters does not bias the
inference on the source parameters as required by the ParSpec.
}
\label{fig:corner_plot}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for the cubic EFT of GR.
We show our results for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
The colors distinguish different $N_{\rm max}$ in the spin expansion and, once more, the vertical
lines mark the 90\% upper credible intervals.
}
\label{fig:cEFT_exec_sum}
\end{figure}
%
\begin{figure}[h]
\includegraphics[width=\columnwidth]{figs/ceft_cdf_varying_threshold_corrected.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_cdf}, but for the cubic EFT of GR.
We see that the CDF curves for $\gm = M_{\rm f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $j = 1$ corrections to
the dominant QNM.
}
\label{fig:cEFT_cdf}
\end{figure}

@@

In this theory, we have considered only $N_{\rm max} = 0$.
%
We find that the addition of spin corrections (while maintaining the same prior
ranges on $\ell_{\rm qEFT}$ as used in the $N_{\rm max} = 0$ study) can result
in waveforms that can have a ringdown segment larger (sometimes seconds long)
than the inspiral-plunge segment in the detectors' frequency band, making the
parameter estimation challenging.
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for the quartic EFT of GR.
We show our results for GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
As there, we mark with the vertical lines the 90\% upper credible intervals.
}
\label{fig:qEFT_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold_corrected.pdf}
\caption{Similar to Fig.~\ref{fig:edgb_cdf}, but for the quartic EFT of GR.
We find that similarly to what happens in dCS gravity and the cubic EFT of GR, we
can place a constrain on $\ell_{\rm qEFT}$ when $\gm = M_{\rm f}$ for both events at 68\% credibility.
}
\label{fig:qEFT_cdf}
\end{figure}

>>>>

\documentclass[twocolumn,
               prd,
               aps,
               superscriptaddress,
               tightenlines,
               nofootinbib,
               eqsecnum,
               amsfonts,
               amsmath,
               longbibliography]{revtex4-1}
\usepackage[flushleft]{threeparttable}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}

@@

\begin{figure}[htb]
\includegraphics[width=\columnwidth]{example_waveform_cubicEFT.pdf}
\caption{(Color online). Gravitational-wave signal with GW150914-like parameters
for both GR (solid line) and cubic EFT of GR (dashed line) with the leading-order
$n=0$ modifications to the fundamental QNM, for $\ell_{\rm cEFT} = 65$~km.
%
The former is computed with the \SEOB{} model, while the latter with the
\pSEOB{} model, with ringdown modifications according to the results in
Table~\ref{tab:ref_theories_qnms}.
%
Top panel: the $+$ polarization $h_{+}(t)$. Bottom panel: the GW amplitude
$|\tilde{h}(t)|$ (left axes) and the instantaneous frequency $f(t)$ (right axes).
}
\label{fig:example_waveform}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{edgb_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function on coupling constant $\ell_{\rm EdGB}$ in EdGB gravity
for GW150914 (top panel), GW200129 (middle panel)
and from combining results (bottom panel).
%
In all panels, different line colors correspond to the inclusion ($N_{\rm max} = 1$) or not ($N_{\rm max} = 0$)
of the linear-in-spin QNM correction.
%
The joint posteriors are shown for illustrative purposes only. As we explain in Fig.~\ref{fig:edgb_cdf} and
in the main text, our analysis of these events fails to satisfy the EFT bound~\eqref{eq:eft_bound}. We mark
with vertical lines the $90\%$ upper credible intervals.}
\label{fig:edgb_exec_sum}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\columnwidth]{edgb_gamma.pdf}
\caption{(Color online).~Posterior distribution function on the dimensionless parameter $\gamma_{\rm EdGB}$,
defined in Eq.~\eqref{eq:def_gamma_gb}. Different line colors distinguish between events, while
different line styles distinguish between different $N_{\rm max}$.
%
We see that this parameter which controls the ParSpec expansion in EdGB gravity
does have maximum support at $\gamma_{\rm EdGB} = 0$. This shows that our model is consistent with GR.
The right panel shows the PDFs in the range $0 \leqslant \gamma_{\rm EdGB} \leqslant 1$. Note how
the curves are flat in this range for both events and when $N_{\rm max} = 1$.}
\label{fig:gamma_gb}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\columnwidth]{edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ for EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{\rm f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines mark the $90\%$ credible levels. Having set the maximum value
of $\varepsilon$ to be 1/2, we see that in no situation the curves pass
through the 0.90 lines. This means that no bound on $\ell_{\rm EdGB}$ can be placed
with the events we analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{dcs_posteriors_combined.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for dCS gravity.
%
We stress that posteriors obtained including $n=1$ corrections, violate conditions~\eqref{eq:parspec_bound}
and therefore should not be used to draw meaningful conclusions. We show them for illustrative purposes
and also to emphasize the importance of taking conditions~\eqref{eq:eft_bound} and~\eqref{eq:parspec_bound}
simultaneously into consideration when analyzing the results of the parameter estimation.
}
\label{fig:dCS_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{dcs_cdf_varying_threshold.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_cdf}, but for dCS gravity.
%
We see that the CDF curves for $\gm = M_{\rm f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $n = 1$ dCS corrections to
the dominant QNM.
}
\label{fig:dcs_cdf}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=0.9\columnwidth]{GW150914_intrinsic_params_remnant.pdf}
\caption{(Color online).~Corner plot showing that the inferred final spin $\chi_{\rm f}$,
remnant mass $M_{\rm f}$ for GW150914, using the same waveform model,
but without (purple contours) and with the non-GR parameters
different from zero (blue contours), for
dCS gravity and $N_{\rm max}=0$.
%
The contours represent 90\% credible levels.
%
We see that the introduction of the non-GR parameters does not bias the
inference on the source parameters as required by the ParSpec.
}
\label{fig:corner_plot}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{ceft_posteriors_combined.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for the cubic EFT of GR.
We show our results for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
The colors distinguish different $N_{\rm max}$ in the spin expansion and, once more, the vertical
lines mark the 90\% upper credible intervals.
}
\label{fig:cEFT_exec_sum}
\end{figure}
%
\begin{figure}[h]
\includegraphics[width=\columnwidth]{ceft_cdf_varying_threshold.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_cdf}, but for the cubic EFT of GR.
We see that the CDF curves for $\gm = M_{\rm f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $j = 1$ corrections to
the dominant QNM.
}
\label{fig:cEFT_cdf}
\end{figure}

@@

In this theory, we have considered only $N_{\rm max} = 0$.
%
We find that the addition of spin corrections (while maintaining the same prior
ranges on $\ell_{\rm qEFT}$ as used in the $N_{\rm max} = 0$ study) can result
in waveforms that can have a ringdown segment larger (sometimes seconds long)
than the inspiral-plunge segment in the detectors' frequency band, making the
parameter estimation challenging.
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for the quartic EFT of GR.
We show our results for GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
As there, we mark with the vertical lines the 90\% upper credible intervals.
}
\label{fig:qEFT_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold.pdf}
\caption{Similar to Fig.~\ref{fig:edgb_cdf}, but for the quartic EFT of GR.
We find that similarly to what happens in dCS gravity and the cubic EFT of GR, we
can place a constrain on $\ell_{\rm qEFT}$ when $\gm = M_{\rm f}$ for both events at 68\% credibility.
}
\label{fig:qEFT_cdf}
\end{figure}
